The yaq Project: \\ A Standard for Scientific Instrumentation

### 2022-11-01 paper.tex
Variety of smaller fixes to spelling and grammar/word choice (largely) suggested by JCW and DDK (#13)

Instrumentation development is a key part of the scientific enterprise.
Novel instruments are typically constructed of many individual components, purchased and home-built.
Orchestration software must communicate with each hardware component in the course of a scientific experiment.
This can involve utilization of many interfaces: NI DAQmx\cite{nidaqmx}, SCPI\cite{scpi}, ModBus\cite{modbus}, PICam\cite{picam}, Thorlabs APT\cite{thorlabs_apt}, among many others.
The challenge of integrating all of these is a frustrating piece of the modern instrument development process.
Weeks can be spent just integrating one new component into an existing project.
In small academic labs, these software interfaces are typically created by student researchers without software development experience.
Student researchers rarely focus on software reusability, and a lack of maintenance and documentation can make such software more difficult to use as time goes on.
Scientists may struggle to rapidly innovate on their experimental design when each hardware addition requires major software development.

@@

We have created a new network-based standard for scientific instrumentation, \yaq{}.
This standard borrows the most important ideas from established projects used by large user facilities while retaining simplicity appropriate for small research labs.
We have built this standard to be self-describing, portable, and reusable wherever possible.
Our primary goal has been to make an interface which simplifies orchestration software development as much as possible.

Here, we discuss the design of the \yaq{} standard in the context of four challenges facing instrument designers.
First, we discuss how particularly challenging hardware interfaces can become seemingly insurmountable barriers to software control.
Next, we discuss how inflexible orchestration software can limit experimental creativity.
Next, we focus on challenges that arise when enhancing or modifying existing instruments with new hardware.
Finally, we discuss the heavy software maintenance burden that many instrument designers face.
In each case, we will highlight how the \yaq{} project is designed to alleviate that challenge.
Several case studies provide a view into the flexible ways that \yaq{} can be applied to perform different scientific experiments.
This paper provides an overview of the concepts and motivations behind \yaq{}.
Refer to the \yaq{} website for a formal specification of the \yaq{} standard.\cite{yaq_fyi}

@@

The first hardware interface barrier: Multiple incompatible interfaces are used to communicate with each component of the system.
\yaq{} provides a unified TCP/IP interface to all hardware peripherals based on the well-described Avro RPC protocol.
The trait system was introduced in pursuit of our primary goal of easing the client development process.
Clients can trust that protocols that implement a given trait will behave in similar ways.
The standardized \yaq{} interface presents the same set of interactions for client-side scientific code, simplifying the experience of using hardware.

>>>>

Instrumentation development is a key part of the scientific enterprise.
Novel instruments are typically constructed of many individual components that are both purchased and home-built.
Orchestration software must communicate with each hardware component in the course of a scientific experiment.
This can involve utilization of many interfaces: NI DAQmx\cite{nidaqmx}, SCPI\cite{scpi}, ModBus\cite{modbus}, PICam\cite{picam}, Thorlabs APT\cite{thorlabs_apt}, among many others.
The challenge of integrating all of these interfaces is a frustrating piece of the modern instrument development process.
Weeks can be spent just integrating one new component into an existing project.
In small academic labs, these software interfaces are typically created by student researchers without software development experience.
Student researchers rarely focus on software reusability, and a lack of maintenance and documentation can make such software more difficult to use as time goes on.
Scientists may struggle to rapidly innovate on their experimental design when each hardware addition requires major software development.

@@

We have created a new network-based communication standard for scientific instrumentation, \yaq{}.
This standard borrows the most important ideas from established projects used by large user facilities while retaining the simplicity appropriate for small research labs.
We have built this standard to be self-describing, portable, and reusable wherever possible.
Our primary goal has been to make an interface which simplifies orchestration software development as much as possible.

Here, we discuss the design of the \yaq{} standard in the context of four challenges facing instrument designers.
First, we discuss how particularly challenging hardware interfaces can become seemingly insurmountable barriers to software control.
Next, we discuss how inflexible orchestration software can limit experimental creativity.
Then, we focus on challenges that arise when enhancing or modifying existing instruments with new hardware.
Finally, we discuss the heavy software maintenance burden that many instrument designers face.
In each case, we will highlight how the \yaq{} project is designed to alleviate that challenge.
Several case studies provide a view into the flexible ways that \yaq{} can be applied to perform different scientific experiments.
This paper provides an overview of the concepts and motivations behind \yaq{}.
Refer to the \yaq{} website for a formal specification of the \yaq{} standard.\cite{yaq_fyi}

@@

The first hardware interface barrier: Multiple incompatible interfaces are used to communicate with each component of the system. % consider making this sentence (and the next paragraph starts) subsection headings 
\yaq{} provides a unified TCP/IP interface to all hardware peripherals based on the well-described Avro RPC protocol.
The trait system was introduced in pursuit of our primary goal of easing the client development process.
Clients can trust that protocols that implement a given trait will behave in similar ways.
The standardized \yaq{} interface presents the same set of interactions for client-side scientific code, simplifying the experience of using hardware.
